---
title: "Plot the placebo data against the election data"
output: html_notebook
---

This notebook runs the main empirical results presented in Figures 7-8. It also performs the visualization in Figure 3 that shows the boost in traffic on election day. 

```{r}
library(plotly)
library(latex2exp)
library(here)
library(tidyverse, warn.conflicts = FALSE)
img_path <- here("img", "nc", "final") 
img_height <- 8
img_width <- 8
theme_set(theme_bw(base_size = 20) + theme(legend.position = "bottom"))
```

> Get input data & specify location of output

```{r}
# input
df <- readRDS(here("processed", "df.Rds"))
monthly <- readRDS(here("processed", "df_monthly.Rds"))

# output
img_path <- here("img", "nc" ,"final") 

# set parameters
election_date <- lubridate::ymd("2018-11-06")
```


Note that we have 44 non-election day weekdays in October and November. As we have only pulled data from October and November, we must drop Oct 1 and Nov 30 since we don't observe adjacent, yielding 42 non-election weekdays. We also drop days adjacent to election day, yielding 40 non-election weekdays.

```{r}
cor_rates <- c()
cor_rates_nw <- c()
cor_counts <- c()
for (date_compare in seq(min(monthly$date), max(monthly$date), by = "1 day")) {
  monthly %>%
    filter(date == date_compare) %>%
    #select(safegraph_place_id, date, sg_count, sg_base_count) %>%
    rename(sg_count_ref = sg_count) %>%
    right_join(select(df, 
                      safegraph_place_id, 
                      perc_over_65,
                      perc_nw),
      by = c("safegraph_place_id")
    )  %>%
    mutate(rate_ref = (sg_count_ref - sg_base_count)/ voters,
           count_ref = (sg_count_ref - sg_base_count)) %>%
    filter(!is.na(rate_ref)) -> df_mod
  cor_rates <- c(cor_rates, cor(df_mod$rate_ref, df_mod$perc_over_65, method = "spearman"))
  cor_rates_nw <- c(cor_rates_nw, cor(df_mod$rate_ref, df_mod$perc_nw, method = "spearman"))
    cor_counts <- c(cor_counts, cor(df_mod$count_ref, df_mod$voters, method = "spearman"))
}
```

# Placebo correlation with rates
## correlation of rate and age
```{r}
tibble(correlations = cor_rates, date = seq(min(monthly$date), max(monthly$date), by = "1 day")) %>%
  mutate(election = if_else(date == election_date, "Election", "Regular")) %>%
  filter(!is.na(correlations)) %>%
  filter(date != election_date + lubridate::days(1),
         date != election_date - lubridate::days(1)) -> marg_cor_age

election_marg_cor_age <- marg_cor_age %>% filter(date == election_date) %>% pull(correlations)
```


## correlation of rate with race
```{r}
tibble(correlations = cor_rates_nw, date = seq(min(monthly$date), max(monthly$date), by = "1 day")) %>%
  mutate(election = if_else(date == election_date, "Election", "Regular")) %>%
  filter(!is.na(correlations)) %>%
  filter(date != election_date + lubridate::days(1),
         date != election_date - lubridate::days(1)) -> marg_cor_race

election_marg_cor_race <- marg_cor_race %>% filter(date == election_date) %>% pull(correlations)

```

# compute p-values
```{r}
marg_pval_age <- nrow(filter(marg_cor_age, 
                             #date!= election_date, 
                             correlations <= election_marg_cor_age))/nrow(marg_cor_age)#nrow(filter(marg_cor_age, date!= election_date))

print(glue::glue("Empirical p-value for age marginal is {marg_pval_age}"))

marg_pval_race <- nrow(filter(marg_cor_race, 
                              #date!= election_date, 
                              correlations <= election_marg_cor_race))/nrow(marg_cor_race)#nrow(filter(marg_cor_race, date!= election_date))

print(glue::glue("Empirical p-value for race marginal is {marg_pval_race}"))
```

# combining the marginal correlations into one plot
```{r}
marg_cor_age %>% 
  mutate(Dem = "age") %>% 
  bind_rows(mutate(marg_cor_race, Dem = "race")) %>%
  ggplot(aes(x= correlations, fill = election)) +
  geom_histogram(binwidth = 0.03, color = "white") +
  theme(legend.title = element_blank()) + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  facet_grid(Dem~.) + 
  xlab(TeX("Rank correlation $\\rho(r(D), r(C(S -Z, V)))$")) + 
    annotate("text", x = -0.13, y = 7.9, size = 5, label = "p-value < 0.05") +   
  ggsave(paste0(img_path, "placebo_cor_combined.pdf"), height = img_height*.8, width = img_width, dpi = 1000)

```

# plot only age (shown in FAcct talk)
```{r}
marg_cor_age %>% 
  mutate(Dem = "age") %>% 
  ggplot(aes(x= correlations, fill = election)) +
  geom_histogram(binwidth = 0.03, color = "white") +
  theme(legend.title = element_blank()) + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  scale_y_continuous(breaks = seq(0,10,2), minor_breaks = seq(0,10,1))  + 
  xlab(TeX("Rank correlation $\\rho")) + 
    annotate("text", x = -0.13, y = 9.9, size = 5, label = "p-value < 0.05") +   
  ggsave(paste0(img_path, "placebo_cor_age.pdf"), height = img_height*.8, width = img_width, dpi = 1000)

```

# placebo correlation on voter count vs sg count
```{r}
tibble(correlations = cor_counts, date = seq(min(monthly$date), max(monthly$date), by = "1 day")) %>%
  filter(!is.na(correlations)) %>%
  mutate(election = if_else(date == election_date, "Election", "Regular")) %>%
  filter(date != election_date + lubridate::days(1),
         date != election_date - lubridate::days(1)) -> corcounts_df # filter out days whose correction was based on voting 
```

```{r}
corcounts_df %>%
  ggplot(aes(x= correlations, fill = election)) +
  geom_histogram(binwidth = 0.05, color = "white") +
  theme(legend.title = element_blank()) + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  xlab(TeX("Rank correlation $\\rho(r(V), r(S -Z))$")) + 
  annotate("text", x = 0.4, y = 5, label = "p-value < 0.05")
  ggsave(paste0(img_path, "placebo_cor_count.pdf"), height = img_height*.8, width = img_width, dpi = 1000)

```




# placebo linear regression 

```{r}
coef_race <- c()
coef_age <- c()
for (date_compare in seq(min(monthly$date) + lubridate::days(1), max(monthly$date) - lubridate::days(1), by = "1 day")) {
  monthly %>%
    filter(date == date_compare) %>%
    rename(sg_count_ref = sg_count) %>%
    right_join(select(df, 
                      safegraph_place_id, 
                      perc_over_65,
                      perc_nw),
      by = c("safegraph_place_id")
    )  %>%
    mutate(rate_ref = (sg_count_ref - sg_base_count)/ voters,
           count_ref = (sg_count_ref - sg_base_count)) %>%
    filter(!is.na(rate_ref)) %>%
    filter(!is.na(perc_nw)) -> df_mod
  if (nrow(df_mod) > 0) {
  mod <- lm(rate_ref ~ perc_nw + perc_over_65, data = df_mod)
  coef_race <- c(coef_race, coef(mod)['perc_nw'])
  coef_age <- c(coef_age, coef(mod)['perc_over_65'])
  } else{
    coef_race <- c(coef_race, NA)
  coef_age <- c(coef_age, NA)
  }
}
```

```{r}
tibble(coef_race = coef_race, coef_age = coef_age,
       date = seq(min(monthly$date) + lubridate::days(1), 
                  max(monthly$date) - lubridate::days(1), 
                  by = "1 day")) %>%
  filter(date != election_date + lubridate::days(1),
         date != election_date - lubridate::days(1)) %>%
  filter(!is.na(coef_age)) %>%
  mutate(election = if_else(date == election_date, "Election", "Regular")) -> coef_df

election_coef_race <- coef_df %>% filter(date == election_date) %>% pull(coef_race)
election_coef_age <- coef_df %>% filter(date == election_date) %>% pull(coef_age)

pval_race <- nrow(filter(coef_df, 
                         coef_race <= election_coef_race))/nrow(coef_df)
pval_age <- nrow(filter(coef_df,  
                        coef_age <= election_coef_age))/nrow(coef_df)

coef_df %>% 
  pivot_longer(cols = coef_race:coef_age, names_prefix = "coef_", names_to = "Demographic", values_to = "Coefficient") -> coef_df

print(glue::glue("Empirical p-value for race is {pval_race}"))

print(glue::glue("Empirical p-value for age is {pval_age}"))
```


# note you will get a warning for the NA values for weekends--this is expected and by design
```{r}
coef_df %>%
  ggplot(aes(x= Coefficient, fill = election)) +
  geom_histogram(binwidth = 0.005, color = "white") +
  theme(legend.title = element_blank()) + 
  scale_fill_viridis_d() + 
  scale_color_viridis_d() + 
  facet_grid(Demographic ~.) + 
  xlab(TeX("Coefficient of linear regression of coverage on age and race")) + 
  annotate("text", x = -0.032, y = 23, size= 5, label = "p-value < 0.05")
  ggsave(paste0(img_path, "placebo_coef.pdf"), height = img_height, width = img_width, dpi = 1000)

```


# Traffic over the month

```{r}
monthly %>%
  group_by(date) %>%
  summarise(sg_counts = sum(sg_count)) %>%
  mutate(x = "All") -> m1
  

monthly %>%
  mutate(sg_count = sg_count - sg_base_count) %>%
  group_by(date) %>%
  summarise(sg_counts = sum(sg_count)) %>%
  mutate(x = "Marginal") -> m2
  
m1 %>%
  bind_rows(m2) %>%
  ggplot(aes(x = date, y = sg_counts, color = factor(lubridate::wday(date, label = TRUE, abbr = FALSE)))) +
  geom_point(size = 4.5) + 
  ylab("SafeGraph traffic") + 
  xlab("Date") + 
  facet_grid(x~.) + 
  #scale_x_discrete(breaks = c(lubridate::ymd("2018-10-01"), lubridate::ymd("2018-11-06"), lubridate::ymd("2018-12-01"))) + 
  theme(legend.title = element_blank()) + 
  geom_vline(xintercept = lubridate::ymd("2018-11-06"), linetype = 2, alpha = 0.35) +
  ggsave(paste0(img_path, "monthly_traffic.pdf"), height = img_height, width = img_width)
```


